---
title: "Telco Churn"
author: "Lanzo Siega"
date: "27/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Checking Working Directory
```{r}
getwd()
```

Assigning "telco" variable
```{r}
telco <- read.csv("WA_Fn-UseC_-Telco-Customer-Churn.csv")
```

Packages
```{r}
install.packages(c("corrplot","tidyverse","dplyr"))

library(dplyr)
library(ggplot2)
library(corrplot)
library(GGally)
```


# Exploratory Statistics

structure of telco
```{r}
str(telco)
View(telco)
summary(telco)
```


replacing "SeniorCitizen" numeric values with "Yes/No" and converting to factor
```{r}
telco$SeniorCitizen <- gsub('0', 'No', telco$SeniorCitizen)
telco$SeniorCitizen <- gsub('1', 'Yes', telco$SeniorCitizen)
telco$SeniorCitizen <- as.factor(telco$SeniorCitizen)

class(telco$SeniorCitizen)
```

Renaming PaymentMethod
```{r}
levels(telco$PaymentMethod) <- c("Bank transfer", "Credit card", "Electronic check", "Mailed check")
```


Replacing NA values in TotalCharges with mean of TotalCharges
```{r}
subset(telco, is.na(telco$TotalCharges))
telco <- na.omit(telco)
sum(is.na(telco))

```

Boxplot for Quantitative Variables
```{r}
boxplot(telco$tenure, main = "tenure")
boxplot(telco$MonthlyCharges, main = "MonthlyCharges")
boxplot(telco$TotalCharges, main = "TotalCharges")
}
```

Correlation Matrix for Quantitative Variables
```{r}
quant <- telco[,c(6,19,20)]

ggcorr(quant,label = T, digits = 3, low = "#ff0f0f", mid = "#ffffff", high = "#0206ed", midpoint = 0)
```

Histogram of Numeric variables grouped by Churn
```{r}


for (i in c(2:length(colnames(telco)))){
  if (is.numeric(telco[,i]) == T ){
     print(ggplot(telco, aes(x = telco[,i], fill = Churn)) + 
            geom_histogram(bins = 10) + labs(x = colnames(telco)[i]))
  }
}

```

Barplot for Qualitative variables
```{r}

for (i in c(2:length(colnames(telco)))){
  if (is.factor(telco[,i]) == T ){
    print(
      ggplot(telco, aes(telco[,i], fill = Churn)) + 
        geom_bar(position = "fill") + 
        labs(x = colnames(telco)[i]) + 
        theme(axis.text.x = element_text(size = 8)))
  }
}

```

Barplot for Churn
```{r}

ggplot(telco, aes(x = Churn, y = (..count../sum(..count..)), fill = Churn)) + 
  geom_bar() + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Frequency")

```

#Classification

Balancing the data (Oversampling, undersampling, SMOTE)
```{r}

install.packages("unbalanced")
library(unbalanced)
levels(telco$Churn) <- c(0,1)

ind_var <- telco[,1:20]
cl <- telco$Churn

usdat <- ubUnder(X=ind_var, Y = cl,perc = 30, method = "percPos")
us_cust <- cbind(usdat$X, usdat$Y)
colnames(us_cust)[21] <- "Churn"
barplot(table(us_cust[,21]), xlab = colnames(us_cust)[21])

osdat <- ubOver(X=ind_var, Y=cl) #best result
os_cust <- cbind(osdat$X, osdat$Y)
colnames(os_cust)[21] <- "Churn"
barplot(table(os_cust[,21]), main = "Balanced Data through Oversampling", ylab = "Count", xlab = colnames(os_cust)[21])

smdat <- ubSMOTE(X=ind_var, Y = cl)
sm_cust <- cbind(smdat$X, smdat$Y)
colnames(sm_cust)[21] <- "Churn"
barplot(table(sm_cust[,21]), xlab = colnames(sm_cust)[21])

```

Algorithms

Splitting into Train/Test sets
```{r}

set.seed(40)
train <- sample(nrow(os_cust), 0.7*nrow(os_cust), replace = FALSE)
TrainSet <- os_cust[train, -1]
TestSet <- os_cust[-train, -1]

summary(TrainSet)
summary(TestSet)

```


Random Forest
```{r}

install.packages("randomForest")
library(randomForest)

rfmod <- randomForest(Churn ~ ., data = TrainSet, importance = T)
rfmod

predTrain <- predict(rfmod, TrainSet, type = "class")
table(predTrain, TrainSet$Churn)

predTest <- predict(rfmod, TestSet, type = "class")
rfacc <- mean(predTest == TestSet$Churn)

rfacc
table(predTest, TestSet$Churn)
```

Naive Bayes
```{r}

install.packages("e1071")
library(e1071)

nbmod <- naiveBayes(Churn ~., data = TrainSet)
nbmod

nbTrain <- predict(nbmod, TrainSet, type = "class")
table(nbTrain, TrainSet$Churn)

nbTest <- predict(nbmod, TestSet, type = "class")
nbacc <- mean(nbTest == TestSet$Churn)
table(nbTest, TestSet$Churn)

```

SVM
```{r}

#also available through the e1071 package

svmod <- svm(Churn ~., data = TrainSet, type = "C-classification", kernel = "linear")
svmod

svmTrain <- predict(svmod,TrainSet, type = "class")
table(svmTrain, TrainSet$Churn)

svmTest <- predict(svmod, TestSet, type = "class")
svmacc <- mean(svmTest == TestSet$Churn)
table(svmTest, TestSet$Churn)

```

Graphing Algorithm Accuracy
```{r}


accuracy <- c(rfacc, nbacc, svmacc)

barplot(accuracy,
        main = "Accuracy of Algorithm",
        xlab = "Algorithm",
        ylab = "Accuracy (%)",
        names.arg = c("Random Forest", "Naive Bayes", "SVM"),
        col = c("#228B22", "#1E90FF", "#CD5C5C"))


```





